### 2016-05-24 project/article_revised_revised/main.tex
final

\documentclass{article}

\usepackage[]{amsmath} 
\usepackage[]{amsthm} 
\usepackage[]{amssymb}

\usepackage[]{thmtools} 
\declaretheorem[style=definition, qed=$\diamondsuit$]{definition}
\declaretheorem[style=definition]{corollary}
\declaretheorem[style=definition, qed=$\spadesuit$]{example}
\declaretheorem[style=definition]{proposition}

\usepackage[]{fontspec} 
\input{code_style.tex}
\usepackage[]{microtype} 
\usepackage[]{enumitem} 
\usepackage[]{graphicx} 
\usepackage[]{subcaption} 
\usepackage[]{makeidx}

\usepackage[margin=2cm]{geometry} 
\makeindex

\usepackage[backend=biber, style=alphabetic]{biblatex} 
\addbibresource{mat2000.bib}
\usepackage[]{url} 
\urlstyle{sf}

\newcommand{\R}{\ensuremath{\mathbb{R}}}
\newcommand{\C}{\ensuremath{\mathbb{C}}}
\newcommand{\A}{\ensuremath{\mathbb{A}}}
\newcommand{\V}{\ensuremath{\mathbb{V}}}

\renewcommand{\P}{\ensuremath{\mathbb{P}}}

\DeclareMathOperator{\Col}{Col}
\DeclareMathOperator{\rank}{rank}
\DeclareMathOperator{\Bl}{Bl}

\usepackage[]{hyperref} 
\usepackage[noabbrev]{cleveref} 
\title{Algebraic Surfaces}
\author{Ivar Haugal{\o}kken Stangeby}
\date{\textsc{MAT2000 - Spring 2016}}

\begin{document}
    \maketitle

    \begin{abstract}
       In this paper we give a brief overview of some of the things one can be
       looking at when studying algebraic surfaces. We use a fairly naive
       approach without the very abstract mathematical machinery that this
       field has to offer. In \cref{sec:preliminaries} we give some of the
       needed mathematical preliminaries. \Cref{sec:intro} motivates the
       definition of an algebraic surface, the main object of study. In
       \cref{sec:singularities} we look at one of the main properties of
       algebraic surfaces, namely the singular points, and proceed by showing
       how to resolve them i \cref{sec:resolving_singularities}. Finally, we
       give an example of how to construct a singular surface with a certain
       symmetry in \cref{sec:constructing_singular_surfaces}.  The images are
       rendered using the software \textsc{Surfer} unless specified otherwise.
       Some auxiliary \textsc{Python} and \textsc{Mathematica} scripts has also
       been used. 
    \end{abstract}

    \tableofcontents

    \section{Preliminaries}
    \label{sec:preliminaries}

    In algebraic geometry the central objects of study is the set of solutions
    to polynomial equations. Finding the zeros of, as an example, the quadratic
    equation
    \begin{equation}
        \notag
        ax^2 + bx + c = 0
    \end{equation}
    is something most people that have been exposed to elementary mathematics
    have done. Later on, one learns how to find the zeroes of more general
    functions, of either one or several variables over some general field of
    coefficients. Of special interest to us, are the polynomials over one or
    several variables, given as elements of a \emph{polynomial ring} over some
    field which we denote $K[x_1, x_2, \ldots, x_n]$.

    As an example, the most familiar polynomial ring is $\R[x]$. Elements of
    this ring include
    \begin{align*}
        f_1(x) = x^2, && f_2(x) = \pi, && f_3(x) = x^{17} + x^{3} - e^{4}, && f_4(x) = x^2 + 1.
    \end{align*} 
    Note that some of these have zeroes in the real numbers $\R$ while some
    does not. Most notably is the function $f_4$ whose zeroes lie in the realm
    of complex numbers, $\C$.

    Given a set of polynomials, $f_1, f_2, \ldots, f_m$ in a polynomial ring
    $K[x_1, x_2, \ldots, x_n]$ we can look at their common zeroes. It is in our
    interest to restrict our attention to polynomial rings where the field of
    coefficients is algebraically closed.

    Given an algebraically closed field $K$, we define \emph{affine
    $n$-space}\index{affine $n$-space} over $K$ to be the set of all $n$-tuples
    of elements of $K$. We denote affine $n$-space by $\A_K^n$ or $\A^n$ if the
    field of coefficients is clear from context.  For our intents and purposes,
    we will be working over the algebraically closed field of complex numbers
    $\C$.

    We now define the mathematical object that is of utmost importance in the
    study of zeroes of polynomials, namely the variety.
    \begin{definition}[Affine algebraic variety {\cite[p.~3]{Har77}}\index{affine algebraic variety}]
        Given a set of polynomials $f_1, f_2, \ldots, f_m$ from a ring of
        polynomials, the set of points $(x_1, x_2, \ldots, x_n)$ in the affine
        space $\A_K^n$ that satisfy
        \begin{equation}
            \notag
            f_i(x_1, x_2, \ldots, x_n) = 0 \quad \text{for } i = 1, 2, \ldots, m
        \end{equation}
        is called an \emph{affine algebraic variety} and is denoted $\V(f_1,
        f_2, \ldots, f_m)$.
    \end{definition}

    Given the polynomial $f(x) = x^2 - 1$, the affine algebraic variety $\V(f)$
    is then the set $\left\{ 1, -1 \right\}$ since $f(1) = f(-1) = 0$. Given an
    additional polynomial $g(x) = x-1$ we have $\V(f, g) = \left\{ 1 \right\}$
    since $g(1) = 0$, but $g(-1) \neq 0$. Finally, if we also consider the
    polynomial $h(x) = x^2$, then $\V(f, g, h) = \emptyset$. We see as a simple
    fact, and state without proof the following.

    \begin{corollary}
        \label{crl:set_one}
        Given a set of polynomials $f_1, f_2, \ldots, f_m$, the variety
        $\V(f_1, f_2, \ldots, f_m)$ can be expressed as a finite intersection
        \begin{equation}
            \notag
            \V(f_1, \ldots, f_m) = \bigcap_{i=1}^{m} \V(f_i).
        \end{equation}
    \end{corollary}
    It is also of interest, especially when aiming to create and visualize
    interesting figures, that the following hold:
    \begin{corollary}
        \label{crl:set_two}
        Given a set of polynomials $f_1, f_2, \ldots, f_m$, the affine algebraic
        variety of the product $f_1f_2\cdots f_m$ is expressible as a finite union
        \begin{equation}
            \notag
            \V(f_1f_2\cdots f_m) = \bigcup_{i=1}^{m} \V(f_i).
        \end{equation}
    \end{corollary}

    With this in mind, we define one of the more interesting properties of
    varieties, and one we will work with a lot in this article, namely the
    singularities. A singular point is, strictly speaking, a point on a surface
    where the behaviour of the surface is slightly strange, say the surface
    passes through itself. We are often interested in studying how a surface
    behaves locally near a point $p$. We do this using \emph{linearization at
    $p$}. This is what we do when we consider the tangent line to a point for a
    single-valued function, the tangent plane for a general surface in three
    dimensions, and a tangent \emph{hyperplane} for general surfaces in $n$
    dimensions. The linearization of a surface given by $f(x_1, x_2, \ldots,
    x_n) = 0$, at a point $p_0$ is given by
    \begin{equation}
        \notag
        L(p) = f(p_0) + J_f(p_0)(p - p_0)
    \end{equation}
    where $J_f(p_0)$ is the Jacobian matrix of $f(x_1, x_2, \ldots, x_n)$
    evaluated at $p_0$. If this matrix has rank less than $n$, then the
    linearization is not uniquely defined, i.e., there are more than one plane
    satisfying the linearization. This leads us to the next definition.

    \begin{definition}[Singularity {\cite[p.~3]{Whi08}}\index{singularity}]
        A singularity of the variety $V = \V(f_1, f_2, \ldots, f_m) \subset
        \A_K^n$ is a point $p \in V$ such that the Jacobian matrix
        \begin{equation}
            \notag
            A = \begin{bmatrix}
                \frac{\partial f_1}{\partial x_1} & \hdots & \frac{\partial f_1}{\partial x_n} \\
                \vdots & \ddots & \vdots \\
                \frac{\partial f_m}{\partial x_1} & \hdots & \frac{\partial f_m}{\partial x_n}
            \end{bmatrix}(p)
        \end{equation}
        has rank strictly less than $\min(m, n)$.
    \end{definition}

    A very useful consequence of this definition deal with the varieties of
    single polynomials and will aid us tremendously in the search for
    singularities.
    \begin{corollary}
        \label{crl:single_poly}
        A point $p$ is a singularity of the variety $V = \V(f) \subset \A_K^n$
        if and only if the partial derivatives of $f$ vanish identically at
        $p$. 
    \end{corollary}
    \begin{proof}
        We recall that the rank of a matrix is the dimension of its column
        space, that is for a matrix $A$, $\rank(A) = \dim(\Col(A))$.  Assume
        first that $p$ is a singularity of $V$. Then $\rank(A) < min(n, m) =
        1$. Consequently, we must have $\rank(A) = 0$, and therefore $A$ equal
        to the zero matrix. All the partial derivatives are therefore zero.
        For the opposite implication, assume that the partials vanish
        identically at $p$. Then the Jacobian matrix reads
        \begin{equation}
            \notag
            A = \begin{bmatrix}
                0 & \hdots & 0
            \end{bmatrix},
        \end{equation}
        and consequently $A$ has rank strictly less than one. Hence $p$ is a
        point of singularity of $V$.
    \end{proof}

    Finally, we want to define the notion of a projective space, as this will
    aid us in the resolution of singularities later on.  Informally, projective
    space generalizes the notion of parallel lines intersecting at infinity.
    The general definition of projective $n$-space is the set of lines in
    $\A^{n+1}$ passing through the origin. For our purposes we will take
    $\A^{n+1} = \C^{n+1}$.

    \begin{definition}[The projective space of dimension $n$\index{projective space}]
        Let $\sim$ be a relation on $\C^{n+1} \setminus \left\{ 0 \right\}$ defined by
        \begin{equation}
            \notag
            (x_1, \ldots, x_{n+1}) \sim (x_1', \ldots, x_{n+1}')
            \iff (x_1', \ldots, x_{n+1}') = \lambda(x_1, \ldots,
            x_{n+1})
        \end{equation}
        where $\lambda$ is some non-zero complex number. This relation can be
        shown to be an equivalence relation. We denote the
        equivalence class of $(x_1, x_2, \ldots, x_{n+1})$ by $\left[ x_1 : x_2
        : \ldots : x_{n+1} \right]$ and define the \emph{projective space of
        dimension $n$} as the set 
        \begin{equation}
            \notag
            \P^{n}_\C = \left\{ \left[ x_1 : \ldots : x_{n+1} \right] \mid (x_1,
            \ldots , x_{n+1}) \in \C^{n+1}\setminus \left\{ 0 \right\}\right\}.
        \end{equation}
        The elements in $\P_\C^n$ are called \emph{homogeneous
        coordinates}\index{homogeneous coordinates}. 
    \end{definition}

    We now have the tools needed to start looking at algebraic surfaces.

    \section{Introduction to algebraic surfaces}
    \label{sec:intro}    
    In this section, we try to motivate the definition of an algebraic surface,
    and look at some examples. Hopefully, this will give an intuitive
    introduction to exactly what we will we looking at. We start by looking at
    real algebraic curves, as a good entry point.

    \subsection{Algebraic Curves}
    \label{sub:algebraic_curves}

    From elementary mathematics, one learns about real valued functions,
    $f(x)$, and how to plot these functions by setting $y = f(x)$ and plotting
    points in the $(x, y)$-plane. This yields a curve of sorts, moving around
    in the plane.  Having an equation $y = f(x)$ we can form the \emph{equation
    of a curve at zero}. We define a new function in two variables and equate
    it to zero. This yields
    \begin{equation}
        \notag
        g(x, y) = y - f(x) = 0.
    \end{equation}
    The set of points along the curve described by this equation is precisely
    the points that lie in the variety $\V(g(x, y))$.  If the function $f(x)$
    is a polynomial in the variable $x$, i.e., $f(x) \in K[x]$, we call the
    function curve of $f(x)$ algebraic. Now, all the curves formed in this way
    corresponds to some function. There are however curves that does not
    correspond to functions and these are given as general polynomials.

    Plotting the result when setting $y^2 - x^3$, the plotted curve does not
    pass the vertical line test and is therefore something different from a
    function graph. Similarity, the equation $y^2 = x^3 + x^2$ defines another
    curve that is not a function graph. These are shown in
    \cref{fig:algebraic_curves}, and are prime examples of curves that exhibit
    a singularity. We will get back to the curve shown in \cref{fig:node}
    later.

    \begin{figure}[h!]
        \centering
        \begin{subfigure}[t]{0.45\textwidth}
            \includegraphics[width=\textwidth]{pictures/node.pdf}
            \caption{The curve given by the equation $y^2 - x^3 - x^2 = 0$.
            This curve is commonly known as ``The Node''. It exhibits a
            singularity at the point $p = (0, 0)$ which can be verified by
            computing the partial derivatives at $p$.}
            \label{fig:node}
        \end{subfigure}        
        ~
        \begin{subfigure}[t]{0.45\textwidth}
            \includegraphics[width=\textwidth]{pictures/cusp.pdf}
            \caption{The curve given by $y^2 - x^3 = 0$, or more commonly known
            as ``The Cusp'', is also an example of a singular algebraic curve.
        The point $p = (0, 0)$ is a point of singularity.}
        \end{subfigure}
        \caption{First examples of algebraic curves. Visualized using
        \textsc{Python}. Code can be seen in \cref{sec:code_snippets}.}
        \label{fig:algebraic_curves}
    \end{figure}

    \subsection{Moving up a dimension}
    \label{sub:moving_up_a_dimension}
    With what we have done so far, we can generalize and move up a dimension.
    Instead of considering curves in the $(x, y)$-plane, we can instead look at
    surfaces in the three-dimensional $(x, y, z)$-space. Following the same
    procedure as above, we define a new function $h(x, y, z) = z - g(x, y)$.
    This function equated at zero yields a surface.  Completely analogous to
    the definition made for curves, we call the surface \emph{algebraic} if
    $h(x, y, z)$ is a polynomial in the variables $x, y, z$. That is $h(x, y,
    z) \in K[x, y, z]$. We give some examples.

    \begin{example}
        \label{exmpl:algebraic_surfaces}
        The following equations satisfy the definition of an algebraic surface:
        \begin{enumerate}
            \item $h_1(x, y, z) = z - xy = 0$,
            \item $h_2(x, y, z) = z - x^2 - y^2 = 0$,
            \item $h_3(x, y, z) = z - (2x^2 - y)(y - x^2) = 0$,
        \end{enumerate}
        while the following does \emph{not}:
        \begin{enumerate}[resume]
            \item $g_1(x, y, z) = \sin(xy) + z = 0$, and
            \item $g_2(x, y, z) = \tan^2(x) + \pi z y$.
        \end{enumerate}
        These examples are visualized in \cref{fig:algebraic_surfaces}.
        \begin{figure}[]
            \centering
            \begin{subfigure}[t]{0.3\textwidth}
                \includegraphics[height=0.6\textwidth]{pictures/example_one.png} 
                \caption{$h_1(x, y, z) = 0$}
            \end{subfigure}
            ~
            \begin{subfigure}[t]{0.3\textwidth}
                \includegraphics[height=0.6\textwidth]{pictures/example_two.png} 
                \caption{$h_2(x, y, z) = 0$}
            \end{subfigure}
            ~
            \begin{subfigure}[t]{0.3\textwidth}
                \includegraphics[height=0.6\textwidth]{pictures/example_three.png} 
                \caption{$h_3(x, y, z) = 0$}
            \end{subfigure}
            \caption{The surfaces that fell under the category algebraic from
            \cref{exmpl:algebraic_surfaces}.}
            \label{fig:algebraic_surfaces}
        \end{figure}
    \end{example}
    We can of course continue this process of introducing new functions and
    move up a dimension for each time indefinitely.  We complete this section
    by formally defining an algebraic surface. In any case, the surfaces are
    defined by the variety of some defining polynomial.
    \begin{definition}[Algebraic surface\index{algebraic surface}]
        Given a set of polynomials $f_1, f_2, \ldots, f_m$ in three variables,
        i.e., in $K[x, y, z]$ we define an \emph{algebraic surface} to be the
        algebraic variety $\V(f_1, f_2, \ldots, f_m)$.  We define the
        \emph{degree}\index{algebraic surface!degree} of an algebraic surface
        to be the highest combined power in the defining polynomials terms, and
        denote this $d$. That is,
        \begin{equation}
            \notag
            d = \max_{\alpha}\left\{ \alpha_1 + \alpha_2 + \ldots \alpha_n \right\}
        \end{equation}
        where $\alpha_i$ is the power of $x_i$ in a term.
    \end{definition}
    We now take a brief look at some of the properties of these surfaces in
    relation to the singularities.

    \section{Singularities}
    \label{sec:singularities}

    When studying algebraic surfaces we typically differentiate between two
    types of surfaces, namely the singular ones and the non-singular ones.
    These two categories are each interesting in their own respect. A lot of
    mathematical theory concerning surfaces only hold for surfaces that exhibit
    no singularities at all. It is therefore often of interest to somehow
    remove these singularities from the surface in order to apply known
    results.

    The theory of singular surfaces on the other hand contains a lot of
    interesting theory that deals with: the classification of singularities,
    what different types of singularities are there; resolution of
    singularities; and deformation theory.

    \subsection{The number of singularities}
    \label{sub:the_number_of_singularities}

    For an algebraic surface of degree $d$ we let
    $\mu(d)$\index{singularity!$\mu(d)$, maximal number} denote the maximal
    number of singularities such a surface can have.  Determining $\mu(d)$ for
    even relatively small $d$ has proven to be surprisingly difficult.
    \cite{Lab14} gives a very intuitive introduction to the world of
    singularity-hunting and the search for the largest possible $\mu(d)$. We
    shall repeat some of the results here. We start with a non-example of a
    singular surface.

    \begin{proposition}[Singularities of the plane]
        Let $V = \V(f(x, y, z)) = \V(ax + by + cz + d)$ be a plane, then $V$ contains no
        singular points.
    \end{proposition} 
    \begin{proof}
        Assume for contradiction that $p \in V$ is a singular point. Then the
        Jacobian matrix of $ax + by + cz + d$ evaluated at $p$ must have rank
        strictly less than 1, or equivalently, by \cref{crl:single_poly}, the
        partial derivatives vanish at $p$. The partial derivatives of our
        defining polynomial is given by
        \begin{align*}
            \notag
            \frac{\partial f}{\partial x} = a, && \frac{\partial f}{\partial y} = b, &&  \frac{\partial f}{\partial z} = c,
        \end{align*}
        however, none of these are dependent on the point we chose. Hence, for
        these partials to vanish at $p$, we need $a = b = c = 0$. However, our
        variety $V$ then reduces to $\V(d)$ where $d$ is some constant. If $d
        \neq 0$, then $\V(d) = \emptyset$, and therefore does not contain $p$
        at all, which contradicts our assumption. If $d = 0$ on the other hand,
        then we are left with the equation $0 = 0$ which does not describe a
        plane. Consequently, $p$ cannot be a singular point of $V$, and the
        plane therefore exhibits no singularities. 
    \end{proof}
    The argument above tells us that for a general algebraic surface of degree
    $d = 1$, we can have no singularities, hence $\mu(1) = 0$.

    We now look at a surface with four singularities.
    \begin{example}[The Cayley cubic\index{the cayley cubic}]
        The Cayley cubic, a surface of degree $d = 3$, is defined by the
        polynomial
        \begin{equation}
            f(x, y, z) = xyz + (1 - x - y - z)(yz + xz + xy).
        \end{equation}
        We are interested in finding the singularities of this surface, if there are any.
        We first compute the partial derivatives and find that
        \begin{align*}
            \frac{\partial f}{\partial x} &= -(y + z)(2x + y + z - 1), \\
            \frac{\partial f}{\partial y} &= -(x + z)(x + 2y + z - 1), \\
            \frac{\partial f}{\partial z} &= -(x + y)(x + y + 2z - 1).
        \end{align*}
        By inspection, we see that the three partials vanish in the point $(1,
        0, 0)$. Due to the symmetric nature\footnote{The study of symmetric
        functions\index{symmetric function} are also of great interest, which
        we will not go into detail here. However, as a brief mention, given a
        function $f(x_1, x_2, \ldots, x_n)$ and a permutation $\sigma \in S_n$,
        where $S_n$ is the symmetry group on $n$ letters, one can consider the
        set $\mathcal{S} = \left\{ \sigma \in S_n \mid f\left(x_{\sigma(1)},
        x_{\sigma(2)}, \ldots, x_{\sigma(n)}) = f(x_1, x_2, \ldots,
        x_n\right)\right\}$ of permutations that fixes the function $f$. For
        the Cayley cubic discussed above, any $\sigma \in S_3$
        fixes the function, and hence the zeros were easily found by symmetry.} of the
        defining equation, we see that $(0, 1, 0)$ and $(0, 0, 1)$ also are zeroes of
        these polynomials.  Finally, the partials also vanish in the point $(0, 0, 0)$.
        We also see that $f(1, 0, 0) = f(0, 1, 0) = f(0, 0, 1) = f(0, 0, 0) = 0$ so
        these are indeed singularities of the Cayley cubic. The symmetries in the
        defining polynomial is seen reflected in the visualized surface shown in
        \cref{fig:cayley}.

        \begin{figure}[h!]
            \centering
            \includegraphics[width=0.3\linewidth]{pictures/cayley_cubic.png}
            \caption{The Cayley cubic, a degree three surface given by the
                equation $f(x, y, z) = xyz + (1 - x  - y - z)(yz + xz + xy)$.
                It is a good example of a surface whose set of singularities
                show a lot of symmetry. It also exhibit the maximal number of
                singularities for a degree $3$ surface, namely 4.}
            \label{fig:cayley}
        \end{figure}
    \end{example}

    From \cite[p.~7]{Lab14} we have the following upper bound on the maximum
    number of singularities for $d \geq 3$,
    \begin{equation}
        \notag
        \mu(d) \leq \frac{1}{2}(d(d-1)^2 - 3)
    \end{equation}
    and plugging in $d = 3$ yields $\mu(3) \leq 4$. However, we have already
    presented a surface of degree $d = 3$ with four singularities, so in fact
    $\mu(d) = 4$. The theory of finding the maximal number of singularities of
    a given degree and constructing surfaces that exhibit all is a very
    technical matter. The thesis \cite{oliver2005hypersurfaces} gives a good
    introduction to the construction and visualization of such surfaces.

    We have now had a brief look at singularities of surfaces, and how to find
    them, we now consider how to remove, or \emph{resolve}, such singularities.

    \section{Resolving singularities}
    \label{sec:resolving_singularities}

    We call a variety that exhibits no singularities \emph{smooth}\index{smooth
    variety}. It is often possible to study a smooth copy of a variety with
    singularities if we move from affine space to projective space. This
    process can be defined in many different ways, with the seemingly most
    prevalent one being through maps from various spaces to each other. In this
    paper we will try to keep it as intuitive as possible, by using the more
    geometrically based procedure.

    When we blow up, we blow up along a \emph{center} of the variety. By making
    sure that this center is contained in the singular locus of the variety, we
    do not change the surrounding surface. We first consider the case where the
    blowup center is a single point, and then an example where the center is a
    line.

    \subsection{Blowing up a point}
    \label{sub:blowing_up_a_point_singularity_}

    We start by examining resolutions of single points. Informally, when we
    blow up a point $p$ in $\A^n$, we replace the point $p$ by an entire copy
    of a projective space $\P^{n-1}$ and try to leave the rest of the space
    $\A^n$ ``unchanged''. We will
    give a few examples, that hopefully will illustrate the geometrical aspect
    of this, but first a formal definition.

    \begin{definition}[Blowup of $\A^{n}$ at a point $p$.]
        We let the \emph{blowup-surface}\index{blowup-surface!in a point}
        $\Bl_p\A^{n}$ be the set of all pairs consisting of a point $p \in
        \A^n$ and a line from $p$ through the origin. The surface is embedded
        in the space $\A^n\times\P^{n-1}$. We write
        \begin{equation}
            \notag
            \Bl_p\A^n = \left\{ (p, L) \mid p \in \A^{n} \text{ and } p \in L \right\}.
        \end{equation}
    \end{definition}
    Now, this is a fairly technical definition, so where do we go from here?
    When blowing up a surface at a point $p$ we are essentially giving the
    surface ``more room'' so that the points creating the singularities can
    pass each other without hassle. In order to aid us with the above
    definition, we show the following claim that tells us under what
    circumstances the point $p$ lie on the line $L$:

    \begin{proposition}
        \label{prop:point_line}
        Given a point $x = (x_1, x_2, \ldots, x_n)$ in $\A^n$ and a point $y =
        [y_1 : y_2 : \ldots : y_n]$ in the space $\P^{n-1}$, the point $x$ is a
        multiple of $y$ if and only if the matrix
        \begin{equation}
            \notag
            A = \begin{bmatrix}
                x_1 & x_2 & \hdots & x_n \\
                y_1 & y_2 & \hdots & y_n
            \end{bmatrix}
        \end{equation}
        has rank less than or equal to 1. Similarly, the rank is less than or
        equal to 1 if and only if each $2\times 2$-minor $x_iy_j - x_jy_i = 0$ for
        $1 \leq i < j \leq n$.
    \end{proposition} 
    \begin{proof}
        In order to verify this, we recall that $\rank(A) = \rank(A^T)$, so we
        chose to work with the transposed matrix instead. Assume that $x$ is a
        multiple of $y$, so $y = \lambda x$ for some scalar $\lambda$. Then
        \begin{equation}
            \notag
            A^T = \begin{bmatrix}
                x_1 \;\, \lambda x_1 \\
                x_2 \;\, \lambda x_2 \\
                \vdots \\
                x_n \;\, \lambda x_n
            \end{bmatrix}
        \end{equation}
        is the associated matrix. If $x = 0$, then $A^T$ is the zero matrix and
        has therefore rank 0. If $x \neq 0$, then we have exactly one pivot
        column in this matrix, hence $\dim(\Col(A^T)) = 1$. In any case,
        $\rank(A) = \rank(A^T) \leq 1$. For the opposite implication. If $A$
        has rank 0, then $A$ is the zero matrix, and hence trivially, $x$ is a
        multiple of $y$, both being zero. If $A$ has rank $1$, then there is
        one linearly independent column in $A^T$, hence the other column can be
        expressed as a multiple of the first. In other words, $x$ is a multiple
        of $y$.

        If there were a non-zero minor $r\times r$ for $r > 1$, then this
        suggests that there are $r$ linearly independent column vectors, but
        this can't occur since $\rank(A^T) \leq 1$.
    \end{proof}

    With the above result, we can rewrite the definition of the blowup-surface
    $\Bl_p\A^n$ as
    \begin{equation}
        \notag
        \Bl_p\A^n = \left\{ (x_1, x_2, \ldots, x_n; y_1 : y_2 : \ldots : y_n)
    \mid x_iy_j = x_jy_i \text{ for } 1 \leq i < j \leq n \right\}
    \end{equation} which is computationally easier to work with.

    We now give some examples of the computations involved when finding
    blowup-surfaces. We start with the resolution of a singularity in an
    algebraic curve.
    \begin{example}[The node]
        The Node, shown in \cref{fig:node}, is a singular algebraic curve with
        a singularity in the point $p = (0, 0)$. The defining polynomial of the
        node is given by
        \begin{equation}
            \label{eq:node}
            f(x, y) = y^2 - x^2 - x^3,
        \end{equation}
        with the curve $\mathcal{C} = \V(f(x, y, z))$. This curve lies in the
        affine plane $\A^2$, while the resulting blowup-surface of
        $\mathcal{C}$ at the point $(0, 0)$ lies in the space $\A^2 \times
        \P^1$. For a point $(x, y) \in \mathcal{C}$ we need to figure out what
        line passes through both $(x, y)$ and the origin $(0, 0)$. However,
        \cref{prop:point_line} tells us exactly under what circumstances this
        occurs. Let $[u : v]$ be a point in $\P^1$. Then, the matrix
        \begin{equation}
            \notag
            A = \begin{bmatrix}
                x & y \\
                u & v
            \end{bmatrix}
        \end{equation}
        has rank less than or equal to one only when $xv - yu = 0$, our only
        $2\times 2$-minor in this case. This gives us the equation $xv = yu$.
        We chose the point $(u, v)$ as a representative for the equivalence
        class $[u : v]$ however, we might as well chose the point $(u, 1)$ or
        $(1, v)$, since we can set either $u = 1$ or $v = 1$ and then multiply
        the representative by $1 / u$ or $1 / v$ respectively. This makes us
        able to eliminate either $u$ or $v$ from $xv = yu$. This gives us two
        different cases:

        \paragraph{Case 1:}
        \label{par:case_1}

        Setting $u = 1$ we get the equation $y = xv$. This can be substituted
        back into \cref{eq:node} in order to eliminate $y$. This gives us a new
        equation, namely
        \begin{equation}
            \notag
            g(x, v) = x^2(v^2 - 1 - x) = 0.
        \end{equation}
        The variety $\V(g(x, v))$ is what we call an \emph{affine
        chart}\index{affine chart} of the blowup-surface $\Bl_p\A^2$. It can be
        visualized in the $(x, v)$-plane.

        \paragraph{Case 2:}
        \label{par:case_2}

        Setting $v = 1$ yields the equation $x = yu$ which makes us able to
        eliminate $x$ from \cref{eq:node}. This gives us a new equation
        \begin{equation}
            \notag
            h(y, u) = y^2(1 - u^2 - u^2y) = 0.
        \end{equation}
        This is the other affine chart of $\Bl_p\A^2$ and can be visualized in
        the $(y, u)$-plane.  The affine charts given by $h(x, y, z)$ and $g(x,
        y, z)$ are shown visualized in \cref{fig:affine_charts_node}.  The
        relation between these two charts, and the original surface can be
        visualized as in \cref{fig:blow_up_intuitive}.

        \begin{figure}[h!]
            \centering
            \begin{subfigure}[t]{0.3\textwidth}
                \includegraphics[width=1.1\textwidth]{pictures/node_affine_1.pdf}
                \caption{The affine chart resulting from setting $u = 1$ and
                eliminating $y$ in the original equation.}
            \end{subfigure}
            ~
            \begin{subfigure}[t]{0.3\textwidth}
                \includegraphics[width=1.1\textwidth]{pictures/node_affine_2.pdf}
                \caption{The affine chart resulting from setting $v = 1$ and
                eliminating $x$ in the original equation}
            \end{subfigure} \\
            \begin{subfigure}[t]{0.3\textwidth}
                \includegraphics[width=1.0\textwidth]{pictures/blow_up_intuitive.pdf}
                \caption{A visual depiction of the blowup surface that clearly
                    shows how the singularity in the origin is resolved. Taken
                from \cite{Whi08}.}
                \label{fig:blow_up_intuitive}
            \end{subfigure}
            \caption{The affine charts of the Node. The vertical line plotted
                in slightly brighter blue, is called the \emph{exceptional
                divisor}\index{exceptional divisor} and corresponds to the
                blowup of what was the origin in the original surface. Note
                that the curves intersect the exceptional divisor in exactly two
                points. This corresponds to the double point singularity in the
                original surface. The top row is visualized in \textsc{Python}.}
            \label{fig:affine_charts_node}
        \end{figure}
    \end{example}

    Blowing up singular surfaces or curves can be great fun and a lot of
    interesting pictures arise. We now take a look at an example where blowing
    up the surface in a single point is not sufficient for removing the
    singularities.

    \subsection{Blowing up a line}
    \label{sub:blowing_up_a_line_of_singularity_}

    In this section we will consider the algebraic surface given by the
    polynomial 
    \begin{equation}
        \label{eq:whitney}
        f(x, y, z) = x^2y - z^2.
    \end{equation}
    The associated variety is $\V(f(x, y, z))$. This surface is commonly known
    as the ``Whitney Umbrella''\index{Whitney umbrella} and is shown visualized in
    \cref{fig:whitney_umbrella}. As we will see shortly, when trying to blow up
    this surface in a single point, say the origin, the resulting surface does
    not seem to be any ``less singular''.

    \begin{figure}[h]
        \centering
        \includegraphics[width=0.3\linewidth]{pictures/blowup_line.png}
        \caption{The ``Whitney Umbrella'' defined by the polynomial $f(x, y, z)
        = x^2y - z^2$. This surface exhibits singularities all along the
        $y$-axis, as can easily be pointed out on the picture.}
        \label{fig:whitney_umbrella}
    \end{figure}

    \begin{example}[Blowup of the Whitney Umbrella in a single point $p$]
        If we were to proceed by trying to blow up the origin, we quickly
        encounter a problem. Following the standard procedure that we used in
        the previous example, we get the defining equations $xs = yr$, $xt =
        zr$ and $yt = zs$ with $r, s$ and $t$ our homogeneous coordinates.
        Contrary to the previous example, we here get three affine charts, one
        for each homogeneous coordinate.

        \paragraph{Case 1:}
        Setting $r = 1$ gives us $y = xs$ and $z = xt$. When plugged into
        \cref{eq:whitney} gives the equation
        \begin{equation}
            \notag
            g(x, s, t) = x^2(xs - t^2) = 0.
        \end{equation}

        \paragraph{Case 2:}
        We start by setting $s = 1$, which gives us $x = yr$ and $z = yt$.
        Eliminating $x$ and $y$ from \cref{eq:whitney} yields the equation
        \begin{equation}
            \notag
            h(y, r, t) = y^2(yr^2 - t^2) = 0.
        \end{equation}

        \paragraph{Case 3:}
        Finally, setting $t = 1$ yields $x = zr$ and $y = zs$, giving an
        equation
        \begin{equation}
            \notag
            i(z, r, s) = z^2(r^2sz - 1) = 0.
        \end{equation}
        These can be visualized, and the surfaces are shown in
        \cref{fig:blowup_line_charts}. As we see in \cref{fig:whitney_blowup} 
        the blowup fails to remove the singularities.

        \begin{figure}[h]
            \centering
            \begin{subfigure}[t]{0.2\textwidth}
                \includegraphics[width=\textwidth]{pictures/blowup_line_2.png}
                \caption{$g(x, s, t) = 0$}
            \end{subfigure}      
            ~
            \begin{subfigure}[t]{0.2\textwidth}
                \includegraphics[width=\textwidth]{pictures/blowup_line_1.png}
                \caption{$h(y, r, t) = 0$}
                \label{fig:whitney_blowup}
            \end{subfigure}      
            ~
            \begin{subfigure}[t]{0.2\textwidth}
                \includegraphics[width=\textwidth]{pictures/blowup_line_3.png}
                \caption{$i(z, r, s) = 0$}
            \end{subfigure}      
        \caption{The affine charts of the blowup of the Whitney Umbrella at the
            point $p = (0, 0, 0)$. As we can see the singularities improve in
            two of the charts, and in one case the chart is smooth. However,
            the middle surface is identical to our original surface, so the
            blow-up has failed to resolve all the singularities of the original
            variety.}
            \label{fig:blowup_line_charts}
        \end{figure}      
    \end{example}

    As seen, in this case it is not sufficient to blow up only in one single
    point. We may have to blow up the surface at a singular line, singular
    curve or in general, any other set. We make a definition of the blowup of a
    space in a line $L$, analogous to the blowup in a point,  We specialize
    this definition for $\A^3$.

    \begin{definition}[Blowup of $\A^3$ at a line $L$] 
        Given a point $p$ and a line $L$, we are interested in the unique plane
        $H$ that contains both $p$ and $L$.  Since it contains $L$, the plane
        can be parametrized by a point $\lambda \in \P^1$.

        We let the blowup-surface $\Bl_L\A^3$ be the set of all pairs
        consisting of a point $p \in \A^3$ and a point $\lambda \in \P^1$ that
        parametrizes the unique plane $H$ that contains both $p$ and $L$.  We
        write
        \begin{equation}
            \notag
            \Bl_L\A^3 = \left\{ (p, \lambda) \in \A^3 \times \P^1 \mid p, L \in H \right\}.
        \end{equation}
    \end{definition}

    Again, this is a fairly technical definition, so where do we go from here?
    Assuming $L$ is parametrized by $x = at, y = bt, z = ct$, with $a, b$ and
    $c$ non-zero, we can form the symmetric equations 
    \begin{equation}
        \notag
        t = \frac{x}{a} = \frac{y}{b} = \frac{z}{c}.
    \end{equation}
    The line $L$ is then completely determined by two of the three possible
    relations $xb - ya = 0$, $xc - za = 0$ and $yc - zb = 0$.  In order to find
    the respective plane, we are interested in knowing when the plane $rx + sy
    + tz = 0$ can be expressed in terms of the line given by $xb - ya = 0$ and
    $yc - zb = 0$. This can be solved as a matrix problem. Given the matrix
    \begin{equation}
        \notag
        A = \begin{bmatrix}
            b & 0 & r \\
            -a & c & s \\
            0 & -b & t
        \end{bmatrix}
    \end{equation}
    one of the columns of $A$ is linearly dependent of the two others if
    $\det(A) = 0$. Solving $\det(A) = 0$ gives the equation
    \begin{equation}
        \notag
        \det(A) = b(ct + sb) + rab = 0.
    \end{equation}
    We can now solve for either one of $r, s$ and $t$. Solving for $r$ we get
    $r = -( ct + sb ) / a$ which we can substitute back into the equation for
    our plane.  The plane is then determined by 
    \begin{equation}
        \notag
        \frac{-ct - sb}{a}x + sy + tz = 0.
    \end{equation}
    Multiplying through by $a$ we achieve
    \begin{align*}
                                t(za - cx) + s(ya - bx) = 0,
    \end{align*}
    which gives us the restrictions that we need. The $\lambda$ that
    parametrizes the plane $H$ is then the point $[t:s] \in \P^1$.

    We illustrate this definition by again considering the Whitney Umbrella.
    \begin{example}[Blowup of the Whitney Umbrella at a line $L$]
        The singularities of the Whitney Umbrella lie along the $y$-axis, which
        is described by the equations $x = z = 0$. Now, our goal here is to
        find under what circumstances a plane $H$ contains the point $p = (x,
        y, z)$ as well as the line $L$. In other words, we want both the
        $y$-axis and $p$ to lie in the plane $rx + sy + tz = 0$.

        We first find the planes that contain the $y$-axis. Choosing the vector
        $(0, 1, 0)$ as a representative for the $y$-axis we compute the cross
        product between this vector and the point $(r, s, t)$ as a
        representative for the plane. This yields a normal vector given by
        \begin{equation}
            \notag
            \vec{n} = \begin{vmatrix}
                \hat{i} & \hat{j} & \hat{k} \\
                r & s & t \\
                0 & 1 & 0
            \end{vmatrix} = (-t, 0, r).
        \end{equation}
        A plane containing the $y$-axis is then given by the equation $-tx + 0y
        + rz = rz - tx = 0$. The point $\lambda$ is in this case the point $[r:
        t] \in \P^1$. We can now consider the two affine charts by setting $r =
        1$ and $t = 1$ respectively.

        \paragraph{Case 1:} 
        \label{par:case_1_}
        Setting $r = 1$ yields $z = tx$ which when plugged into $x^2y - z^2 = 0$
        gives us $x^2y - (t^2x^2) = x^2(y - t^2) = 0$. This affine chart can be
        visualized in the $(x, y, t)$-plane.

        \paragraph{Case 2:}
        \label{par:case_2_}
        Setting $t = 1$ yields $x = rz$ which when plugged into $x^2y-z^2 = 0$
        gives the equation $r^2z^2y - z^2 = r^2(z^2y - 1) = 0$. This can be
        visualized in the $(r, y, z)$-plane.

        The two affine charts are shown visualized in \cref{fig:affine_line}.

        \begin{figure}[h!]
            \centering
            \begin{subfigure}[t]{0.3\textwidth}
                \includegraphics[width=\textwidth]{pictures/line_blowup_affine_1.png}
                \caption{The surface given by the equation $x^2(y - t^2) = 0$. The
                    curve of intersection between this surface and the exceptional
                    divisor looks like it can be ``pinched'' into a double line, i.e., the
                original line of singularity.}
            \end{subfigure}
            ~
            \begin{subfigure}[t]{0.3\textwidth}
                \includegraphics[width=\textwidth]{pictures/line_blowup_affine_2.png} 
                \caption{The surface given by the equation $r^2(z^2y - 1) = 0$. The
                    intersection between this surface and the exceptional divisor can
                    also be ``fused'' together to form a double line, by laying the
                surfaces flat on top of each other.}
            \end{subfigure}
            \\
            \begin{subfigure}[t]{0.3\textwidth}
                \includegraphics[width=\textwidth]{pictures/affine_1_exceptional.pdf} 
            \end{subfigure}
            ~
            \begin{subfigure}[t]{0.3\textwidth}
                \includegraphics[width=\textwidth]{pictures/affine_2_exceptional.pdf} 
            \end{subfigure}
            \caption{The two affine charts of the variety $\V(x^2y-z^2)$. The
                top row is rendered by \textsc{Surfer}, and the bottom row
                rendered by \textsc{Mathematica.} The exceptional divisor in
                the top row has been divided out and is therefore not shown
                because \textsc{Surfer} does a poor job at rendering planes. As
                seen, the charts are smooth and exhibits no singularities. The
                intersection with the exceptional divisor is rendered in blue
                in the bottom row. The bottom row is visualized using
            \textsc{Mathematica}.}
            \label{fig:affine_line}
        \end{figure}
    \end{example}

    \section{Constructing singular surfaces}
    \label{sec:constructing_singular_surfaces}

    In this section we take a brief look at how one can construct singular
    surfaces with a certain symmetry. The way we will be doing this is based on
    the set theoretic results given in \cref{crl:set_one} and
    \cref{crl:set_two}. This is purely exploratory and is merely a matter of
    trying things out and seeing what the results are.

    As a baseline for our surface, we wish to use the zero set of the function
    \begin{equation*}
        f(x, y, z) = (x - y)(x + y)(x - z)(x + z)(y - z)(y + z).
    \end{equation*}
    We let $\mathcal{B} = \V(f(x, y, z))$ This surface is the intersection of
    six planes, as shown in \cref{fig:construction_one}. This surface has a
    very nice set of singularities, namely on the intersection between the
    planes. It also has a nice symmetry that we want our final surface to
    inherit.

    We can now take $\mathcal{B}$ and ``flesh it out'' by adding another
    polynomial to the defining function. We want to add a surface that has a
    group of symmetry that contains the symmetry group of $\mathcal{B}$. In our
    case, the unit sphere seems a reasonable choice. The unit sphere is given
    by $g(x, y, z) = x^2 + y^2 + z^2 - 1$. We let $\mathcal{S} = \V(g(x, y,
    z))$. Define $\mathcal{B} + \mathcal{S} = \V(f(x, y, z) + g(x, y, z))$.
    This surface has now gone from singular to smooth, but, while a bit hard to
    see, it does retain all the symmetries of $\mathcal{B}$ as shown in
    \cref{fig:construction_two}.

    In order to regain the set of singularities, we can instead add powers of
    $\mathcal{S}$ to $\mathcal{B}$. By powers of $\mathcal{S}$ we mean
    $\mathcal{S}^n = \V(g(x, y, z)^n)$. If $n$ is an even number, we get a
    surface with holes as shown in \cref{fig:construction_three}. If we instead
    were to take $n$ an odd number, we get a solid surface with no holes, as in
    \cref{fig:construction_four}. The higher the value we chose for $n$, the
    more protruding the nodes of the surface become.

    \begin{figure}[h]
        \centering
        \begin{subfigure}[t]{0.2\textwidth}
            \includegraphics[width=\textwidth]{pictures/construction_one.png}
            \caption{$\mathcal{B}$}
            \label{fig:construction_one}
        \end{subfigure}
        ~
        \begin{subfigure}[t]{0.2\textwidth}
            \includegraphics[width=\textwidth]{pictures/construction_1_5.png}
            \caption{$\mathcal{B}+ \mathcal{S}$}
            \label{fig:construction_two}
        \end{subfigure}
        ~
        \begin{subfigure}[t]{0.2\textwidth}
            \includegraphics[width=\textwidth]{pictures/construction_two.png}
            \caption{$\mathcal{B} + \mathcal{S}^n$ for $n$ even.}
            \label{fig:construction_three}
        \end{subfigure}
        ~
        \begin{subfigure}[t]{0.2\textwidth}
            \includegraphics[width=\textwidth]{pictures/construction_odd.png}
            \caption{$\mathcal{B} + \mathcal{S}^n$ for $n$ odd.}
            \label{fig:construction_four}
        \end{subfigure}
        \caption{Constructing a singular surface with a given symmetry and set
            of singularities. The surface $\mathcal{B}$ is the product of six
            planes. The surface $\mathcal{S}$ is the unit sphere. By adding
            various powers of $\mathcal{S}$ to $\mathcal{B}$ we achieve a
        surface that retains the symmetries and singularities of
    $\mathcal{B}$.}
        \label{fig:construction}
    \end{figure}

    \section{Conclusion}
    \label{sec:conclusion}

    This paper was meant as a way of introducing the reader to some of the
    aspects of algebraic geometry, specifically dealing with algebraic
    surfaces. The intention was to, by restricting our attention to specific
    examples, avoid having to take into account the fairly theoretical
    mathematics that lie behind the scenes.  We have briefly touched upon
    different kinds of theory related to these algebraic surfaces, which might
    have piqued an interest.  Algebraic surfaces, while being very intuitive
    objects in some sense, trying to figure out what they really are can
    quickly prove a challenge because the mathematics is so involved.

    Algebraic surfaces can also work as a solid entry point into visualization
    techniques and computational mathematics, and while not talked about here,
    we have seen some end results of the mathematical endeavours related to this
    area of expertise, through the usage of \textsc{Surfer},
    \textsc{Mathematica} and \textsc{Python}.

    As the author of this text, it has been great fun trying to wrap my head
    around these things, and I hope I will get the opportunity to explore this
    field of mathematics in greater detail.  
    \clearpage \appendix

    \section{Installing visualization software}

    The algebraic surface visualization industry is definitely not a million
    dollar business, hence there are only a select few ``out of the box''
    visualization softwares out there. Most notably are \textsc{Surf},
    \textsc{Surf-Ex}, \textsc{Spicy} and \textsc{Surfer}. While fairly modest
    in its features, \textsc{Surfer} is the one we have chosen for this paper,
    due to its ease of installation and ease of use. In that regard, it is
    perfect for this kind of introductory/exploratory text.  The software can
    be downloaded from
    \begin{center}
        \url{https://imaginary.org/program/surfer#all-downloads}
    \end{center}
    and is compiled for use with Windows, \textsc{OS-X} and Linux operating
    systems. Debian based Linux distributions will want to download the
    \textsf{.deb} package, where as Red-Hat systems will want to use the
    \textsf{.rpm} package.

    If one want more options, then \textsc{Surf-Ex} is the software of choice,
    however this has proven to be very difficult to install as it uses
    \textsc{Surf} as a back end. In any case, the software can be found on
    \begin{center}
        \url{http://surfex.algebraicsurface.net} 
    \end{center}
    along with installation guidelines.

    \section{Code snippets}
    \label{sec:code_snippets}

    \lstset{deletekeywords={xrange}}
    \begin{TMcode}{python}{curves.py -
        \url{http://stackoverflow.com/a/2484594}}{Plotting algebraic curves}
    import matplotlib.pyplot
    from numpy import arange
    from numpy import meshgrid

    delta = 0.025
    xrange = arange(-5, 5.0, delta)
    yrange = arange(-2.5, 2.5, delta)
    X, Y = meshgrid(xrange,yrange)

    # F is one side of the equation, G is the other
    F = X**2*(1 - Y**2 - X*Y**3)
    G = 0

    matplotlib.pyplot.axis('on')
    matplotlib.pyplot.grid('on')
    matplotlib.pyplot.contour(X, Y, (F - G), [0])

    matplotlib.pyplot.savefig('output.pdf')
    \end{TMcode}

    \clearpage
    \begin{TMcode}{mathematica}{intersection.nb - \url{http://community.wolfram.com/groups/-/m/t/178218?p_p_auth=lT7er3Rd}}{Plotting the intersection between to surfaces}

    (* affine chart and exceptional divisor *)
    h = z^2 y - 1;
    g = x^2;

    ContourPlot3D[
        {h == 0, g == 0}, 
        (* bounding box dimensions *)
        {x, -3, 3}, 
        {y, -3, 3}, 
        {z, -3, 3}, 
        MeshFunctions -> {
            Function[ {x, y, z, f}, h - g ]
        }, 
        MeshStyle -> {
            { Thick, Blue }
        }, 
        Mesh -> {{0}}, 
        ContourStyle -> Directive[
                Green, Opacity[0.5], Specularity[White, 30]
        ], 
        PlotPoints -> 60,
        SphericalRegion -> True
    ]
    \end{TMcode}

    \printbibliography 
\end{document}